Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-09-13 paper/intro.tex
Revise intro to use macros

Formal
verification that the transformed system model meets its requirements
is accomplished via AGREE~\cite{agree2013} (Assume Guarantee Reasoning
Environment).
%The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
AGREE is a {\em compositional assume-guarantee} style model checker
for AADL models that attempts to prove properties about one layer of
an architecture using properties allocated to its subcomponents.
Cyber-resilient code implementing the verified model is then
automatically generated using the High Assurance Modeling and Rapid
Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If
desired, this code can be targeted to the formally verified seL4
secure microkernel~\cite{sel4-2009}.

@@

Two cyber-resiliency transformations are discussed: (1) the insertion
of a filter to prevent malformed data from a malicious actor from
being propagated to downstream components, and (2) the insertion of a
monitor to detect (and alert) unexpected behaviors arising from
untrusted components. These transformations not only change the
architecture of the model by adding in new components; they also
generate a formal specification in the form of a \emph{code contract} in the AGREE language for their intended behavior. 
A code contract describes the computation of a high-assurance component and not just what it computes.
These code contracts
are sufficient for model checking to prove that---due to the newly
included high-assurance components---the hardened system meets its
cyber-resiliency requirements.

Another novel aspect of the approach is the synthesis of the AGREE
code contracts for the high-assurance components to CakeML, a verified
compiler implementation for the functional programming language
ML \cite{cakeml}. This paper describes in detail the synthesis path
from code contracts to CakeML code, providing a formal framework in
which to argue correctness. CakeML then provides a verified
compilation path to several different target binaries (and also
proving that the meaning of the CakeML source code is exactly
preserved in the final binaries). Assuming that the execution schedule
of the deployed cyber-hardened system is as intended by the AADL
model, and that the HAMR-generated communication fabric delivers
messages between components as expected, the AGREE model checking
results then hold for the deployed system, \ie, the system will detect
and prevent the indicated cyber-vulnerabilities over all possible
finite inputs. Preliminary work has shown how to lift this result to
infinite input traces as these systems are inherently reactive and
intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

@@

The approach is illustrated by a simple example in
\secref{sec:example}. AGREE verification and specification are
presented in \secref{sec:agree} and \secref{agree-semantics}.
Code contracts are defined in \secref{sec:code-contracts} with a method
for testing code contracts in the AGREE framework before synthesis given
\secref{sec:testing}.
The synthesis pathway is covered in Section~\ref{sec:synthesis}. A
case study applying these transformations to an Unmanned Aerial
Vehicle (UAV) system that uses the Air Force Research Laboratory's
OpenUxAS services for route planning is presented in
Section~\ref{sec:case-study}. Here the transforms add filters to guard
against malformed input and monitors to guard against ground station
spoofing and malicious flight plans from OpenUxAS. The case study
system is significantly more complex than the simple example and shows
the viability of the modeling approach to a full-scale industrial
design.

The BriefCASE tools are open source and publicly
available \cite{fmide}, as is the motivating example \cite{repo}, and
the full UxAS model with its deployment in seL4 \cite{phase2, camkes}.
Videos demonstrating the use of the BriefCASE tools to build the UAV
example presented in Section~\ref{sec:case-study} are also
available \cite{case}.

>>>>

Formal
verification that the transformed system model meets its requirements
is accomplished via \agr~\cite{agree2013} (Assume Guarantee Reasoning
Environment).
%The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
\agr\ is a {\em compositional assume-guarantee} style model checker
for AADL models that attempts to prove properties about one layer of
an architecture using properties allocated to its subcomponents.
Cyber-resilient code implementing the verified model is then
automatically generated using the High Assurance Modeling and Rapid
Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If
desired, this code can be targeted to the formally verified seL4
secure microkernel~\cite{sel4-2009}.

@@

Two cyber-resiliency transformations are discussed: (1) the insertion
of a filter to prevent malformed data from a malicious actor from
being propagated to downstream components, and (2) the insertion of a
monitor to detect (and alert) unexpected behaviors arising from
untrusted components. These transformations not only change the
architecture of the model by adding in new components; they also
generate a formal specification in the form of a \emph{code contract} in the \agr\ language for their intended behavior. 
A code contract describes the computation of a high-assurance component and not just what it computes.
These code contracts
are sufficient for model checking to prove that---due to the newly
included high-assurance components---the hardened system meets its
cyber-resiliency requirements.

Another novel aspect of the approach is the synthesis of the \agr\ 
code contracts for the high-assurance components to \ckml, a verified
compiler implementation for the functional programming language
ML \cite{cakeml}. This paper describes in detail the synthesis path
from code contracts to \ckml\ code, providing a formal framework in
which to argue correctness. \ckml\ then provides a verified
compilation path to several different target binaries (and also
proving that the meaning of the \ckml\ source code is exactly
preserved in the final binaries). Assuming that the execution schedule
of the deployed cyber-hardened system is as intended by the AADL
model, and that the HAMR-generated communication fabric delivers
messages between components as expected, the \agr\ model checking
results then hold for the deployed system, \ie, the system will detect
and prevent the indicated cyber-vulnerabilities over all possible
finite inputs. Preliminary work has shown how to lift this result to
infinite input traces as these systems are inherently reactive and
intended to run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

@@

The approach is illustrated by a simple example in
\secref{sec:example}. \agr\ verification and specification are
presented in \secref{sec:agree} and \secref{agree-semantics}.
Code contracts are defined in \secref{sec:code-contracts} with a method
for testing code contracts in the \agr\ framework before synthesis given
\secref{sec:testing}.
The synthesis pathway is covered in Section~\ref{sec:synthesis}. A
case study applying these transformations to an Unmanned Aerial
Vehicle (UAV) system that uses the Air Force Research Laboratory's
OpenUxAS services for route planning is presented in
Section~\ref{sec:case-study}. Here the transforms add filters to guard
against malformed input and monitors to guard against ground station
spoofing and malicious flight plans from OpenUxAS. The case study
system is significantly more complex than the simple example and shows
the viability of the modeling approach to a full-scale industrial
design.

The \brfcs\ tools are open source and publicly
available \cite{fmide}, as is the motivating example \cite{repo}, and
the full UxAS model with its deployment in seL4 \cite{phase2, camkes}.
Videos demonstrating the use of the \brfcs\ tools to build the UAV
example presented in Section~\ref{sec:case-study} are also
available \cite{case}.
